Array

### 2014-01-03 DOCUMENTATION/AdvancedTypes.tex
Update the example and rule to support array.

\usepackage[fleqn]{amsmath}
\usepackage{xspace,framed,here,keystroke}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}
\usepackage{minted}   %Used to syntax highlight

@@

\definecolor{bg}{rgb}{0.95,0.95,0.95}

@@

The B0 language supports attribution directly between arrays or structure, even though the attribution
must have the same domain type and range type. \VGM{Project Decision: How translate a direct 
attribution between advanced types, for example, arrayA := arrayB ? The Commenc 
uses a function called ``memmove''.  %TODO to add this restriction in become equals.
 }

\subsection{Relation arrays B0 and LLVM-IR}
\VGM{It must be extended to support structure.}

@@

The essential elements of array are its name, size and type. The
following code declares an array global called ``arr'' with 100 integer
numbers with 32 bits and initialized with zero.

\begin{footnotesize}
\begin{minted}[frame=single,linenos=true,bgcolor=bg]{llvm}
@arr = internal global [100 x i32] zeroinitializer
\end{minted}
\end{footnotesize}

@@

\begin{footnotesize}
\begin{minted}[frame=single,linenos=true,bgcolor=bg]{llvm}
%4 = getelementptr  ([100 x i32]* @Array__arr, i64 0, i64 %3)
\end{minted}
\end{footnotesize}

The first parameter defines the type and name of the variable, and the others are
indexes to select the element in the data type. The second parameter (i64 0)
selects the first element in the pointer (* @Array\_\_arr). The last element,
(i64 \%3) selects the value in position obtained from \%3 of the array of 32-bit
integer values. When the array is multidimensional, the news index must be added
separated by comma to get the specific element.

@@

\subsubsection{Setting an element to array}

The following blocks illustrate a B code setting and getting values with array
``arr''  and its translation to LLVM-IR. 
\begin{footnotesize}

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}
  arr(2):=5;
  i = arr(2);
\end{minted}
\end{footnotesize}

\begin{footnotesize}
\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{llvm}
  store i32 5, i32* getelementptr  ([100 x i32]* @arr, i32 0, i64 2)
  %1 = load i32* getelementptr  ([100 x i32]* @arr, i32 0, i64 2)
  store i32 %1, i32* @i, align 4
\end{minted}
\end{footnotesize}

TODO: show a simple example getting or setting an element of array 
multidimensinal using the rule $ArrayItem$. For example, show the translation of 
this code: $ arr(2,3) :=4$.

@@

\paragraph{Array item} selects an item in expressions as ``$a(n1,...,n_{n})$'', 
where $a$ is $n.base$ and $n1,...,n_{n}$ is a sequence of indices that 
calculates the address of element.

\begin{align*}
\lefteqn{\trad{n}{ArrayItem} \isdef} \\
&  \LET p_1 \sep v_1 \sep t_1  = \trad{n.base}{Exp}  \IN \\
&  \LET p_2 \sep v_2 \sep t_2  = \trad{n.indices}{\ListOf{Exp}}  \IN \\
&  \LET v =  \variable \AND  t = \trad{\B{n.type}}{Type}  \IN \\
&  \llvm{"\PH{p_1} \PH{p_2} \PH{v} =  getelementptr \PH{t_1}, \PH{v_1}, i32 0, i32  \PH{v_2}\nl"}  \sep v \sep t\\
\end{align*}

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}
 array1 := array2;
 structure1 := structure2;
\end{minted}

>>>>

\usepackage[fleqn]{amsmath}
\usepackage{xspace,framed,here,keystroke}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}

\usepackage{minted}   %Used to syntax highlight
\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newminted{llvm}{frame=single,firstnumber=1,linenos=true,fontsize=\scriptsize,bgcolor=bg}
\newminted{pascal}{frame=single,firstnumber=1,linenos=true,fontsize=\scriptsize,bgcolor=bg}
\newminted{c}{frame=single,firstnumber=1,linenos=true,fontsize=\scriptsize,bgcolor=bg}

@@

@@

The B0 language supports attribution directly between arrays or structure, even though the attribution
must have the same domain type and range type.

\subsection{Relation arrays B0 and LLVM-IR}

@@

The essential elements of array are its name, size and type. The
following code declares an array global called ``Array3'' with 6000 (10*20*30)
integer numbers with 32 bits and initialized with zero.

\begin{llvmcode}
@Array3 = internal global [10 x [20 x [30 x i32]]] zeroinitializer, align 16
\end{llvmcode}

@@

\begin{llvmcode}
  %7 = getelementptr [10 x [20 x [30 x i32]]]* @Array3, i32 0, i64 2, i64 3, i64 4
\end{llvmcode}

The first parameter defines the type and name of the variable, and the others are
indexes to select the element in the data type. The second parameter (i32 0)
selects the first element in the pointer (* @Array3). The last elements 
(i64 2, i64 3, i64 4) selects the value from position (2,3,4) storing it in \%7.
This array is multidimensional then several index must be declared separated
by comma to get the specific element.

@@

\subsubsection{Setting and getting array}

The following blocks illustrate a C code setting a value to array
and its translation to LLVM-IR.

\begin{ccode} 
  int x=2; int y=3 ; int z=4 ;
  Array3[x][y][z]=7;
\end{ccode}

This example is translated in four steps. First, the variables \%x, \%y, \%z 
are allocated. Second, the respective values 2, 3, 4 are stored in variables. 
Third, the variables are loaded to local variables and converted to 64 bits. 
Finally, the address of array is calculated in local variable \%7 and the value 
7 is stored in it.

\begin{llvmcode}
%x = alloca i32, align 4
%y = alloca i32, align 4
%z = alloca i32, align 4
store i32 2, i32* %x, align 4
store i32 3, i32* %y, align 4
store i32 4, i32* %z, align 4
%1 = load i32* %z, align 4
%2 = sext i32 %1 to i64
%3 = load i32* %y, align 4
%4 = sext i32 %3 to i64
%5 = load i32* %x, align 4
%6 = sext i32 %5 to i64
%7 = getelementptr[10x[20x[30xi32]]]*@Array3,i32 0,i64 %6,i64 %4,i64 %2;getting an element
store i32 7, i32* %7, align 4;setting an element
\end{llvmcode}

@@

\paragraph{Array item} selects an item in expressions as ``$a(n1,...,n_{n})$'', 
where $a$ is $n.base$ and $n1,...,n_{n}$ is a sequence of indices that 
calculates the address of element. This rule also supports to translate 
expressions of arrays.

\begin{align*}
\lefteqn{\trad{n}{ArrayItem} \isdef} \\
&  \LET p_1 \sep v_1 \sep t_1  = \trad{n.base}{Exp}  \IN \\
&  \LET p_2 \sep v_2 \sep t_2  = \trad{n.indices}{\ListOf{Exp}}  \IN \\
&  \LET v =  \variable \AND  t = \trad{\B{n.type}}{Type}  \IN \\
&  \llvm{"\PH{p_1} \PH{p_2} \PH{v} =  getelementptr \PH{t_1}, \PH{v_1}, i32 0 \ListOf{(,\PH{t_2} \PH{v_2})}\nl"}  \sep v \sep t\\
\end{align*}

@@

\begin{pascalcode}
 array1 := array2;
 structure1 := structure2;
\end{pascalcode}
